---
title: "plotting"
output: html_document
---

# ITM and elevation -- preliminary study.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(leaflet)
library(dplyr)
library(forcats)
library(viridis)
library(lme4)
library(lmerTest)
library(ggforce)
library(plotly)
library(sjPlot)
library(ggeffects)
library(spatstat)
library(ggraph)
library(igraph)
library(ggrepel)
library(ggmap)
library(geosphere)
library(phangorn)
library(ggtda)
```


## Descriptive stats 

### Average elevation across different languages

```{r}
data <- read.csv('elevation_villages_correct.csv')
data
```

```{r}
data %>% 
  group_by(language) %>% 
  slice(which.min(elevation)) %>% 
  select(eng_vil_name, elevation) %>% 
  write.csv('minimum_vil.csv')

```

```{r}
ggplot(data=data, aes(x=fct_reorder(language, elevation, .fun=mean, .desc=TRUE), y=elevation))+
  theme_minimal()+
  # stat_summary(fun.data = mean_se,  
  #                geom = "errorbar", color='black', alpha=0.3)+
  geom_jitter(color='grey', alpha=0.1)+
  stat_summary(aes(color=Family), fun = "mean", geom = "point", size=2)+
  labs(y = "Elevation", x='Language', color='Family')+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = "top")+
  scale_color_viridis(discrete = TRUE, option = "D")+
  coord_flip()+
  theme(axis.text.x=element_text(angle=0, hjust=0.5),
        # text = element_text(family = "Andale Mono"),
        legend.position = 'bottom')+
  geom_hline(aes(yintercept=mean(elevation)), color='red', linetype='dotted')+
  guides(colour = guide_legend(nrow = 1))

ggsave('elev_lang.jpg', width = 8, height = 5)
```

```{r}
data$log_c <- 0
data$log_c <- log(data$census_1926)

data$el_100 <- plyr::round_any(data$elevation, 1, f = ceiling)

data %>%
  group_by(el_100) %>%
  summarise(mean_c = mean(log_c)) %>%
  ggplot(aes(x=el_100, y=mean_c))+
    # geom_smooth()+
    geom_point()
```


### Average elevation across different languages (groupped by family)

```{r}
ggplot(data=data, aes(x=fct_reorder(language, elevation, .fun=mean, .desc=TRUE), y=elevation))+
  theme_bw()+
  # geom_boxplot(aes(fill=Family))+
  geom_dotplot(aes(fill=Family, color=Family), binaxis = "y", stackdir = "center", position = "dodge", binwidth = 60, method = 'histodot')+
  facet_wrap(vars(Family), scales = "free_y", strip.position = 'top')+
  labs(y = "Elevation", x='Language', size=' 1926 Census\n data')+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = "top")+
  scale_color_viridis(discrete = TRUE, option = "D")+
  coord_flip()+
  theme(legend.position = "none",
        legend.key = element_blank(), 
        strip.text = element_text(face="bold"), 
        axis.text.x=element_text(angle=0, hjust=0.5)) 

ggsave('elev_lang.jpg', width = 8, height = 5)
```


```{r}
data[data$language == 'Godoberi ',] %>% 
  write.csv('godoberi.csv')
```


### Average elevation across families (groupped by family)

```{r}
ggplot(data, aes(y=elevation, x=Family))+
  theme_bw()+
  geom_dotplot(aes(fill=Family, color=Family), binaxis = "y", stackdir = "center", position = "dodge", binwidth = 41, method = 'histodot')+
  theme(legend.position = "none",
        legend.key = element_blank(), 
        strip.text = element_text(face="bold"), 
        axis.text.x=element_text(angle=0, hjust=0.5))+
  # stat_summary(fun = "mean", geom = "crossbar", size=0.5, alpha=0.1, color='grey')+
  # geom_hline(aes(yintercept=mean(elevation)), color='red', linetype='dotted')

ggsave('family_dist.jpg', width = 8, height = 5)
```



```{r}
# dplyr::select(data, elevation.round, census_1926) %>% 
#   group_by(elevation.round) %>% 
#   summarise(mean_pop = mean(census_1926), count=n()) %>%
#   ggplot(aes(x=elevation.round, y=mean_pop))+
#   geom_point()
```


## Generalized Linear Models


```{r}
all <- read.csv('all.csv')
```

### Average elevation across different number of L2's

```{r}
ggplot(data = all, aes(y=elevation, x=factor(number.of.lang.strat)))+
  theme_bw()+
  geom_violin(fill='grey', color='grey', trim = TRUE)
```


```{r}
# events <- 0:4
# density <- dpois(x = events, lambda = mean(all$number.of.lang.strat))
# 
# ggplot()+
#   geom_point(aes(y=density, x=0:4, color='Poisson distribution'))+
#   geom_line(aes(y=density, x=0:4, color='Poisson distribution'))+
#   ylim(0, 1)+
#   geom_point(aes(x=co$x, y=co$freq/sum(co$freq), color='Actual data'))+
#   geom_line(aes(x=co$x, y=co$freq/sum(co$freq), color='Actual data'))+
#   labs(y = "Probability", x='ITM')
```


### Poisson and linear regressions

```{r}
fit_p <- glm(number.of.lang.strat ~ elevation + year_of_birth*sex + language.population*residence, family="poisson", data=all)
```

```{r}
fit_lin <- glm(number.of.lang.strat ~ elevation + year_of_birth*sex + language.population*residence, data=all)
```

Linear: 

```{r}
ggplot(ggpredict(fit_lin, terms = c('elevation')), aes(x, predicted)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1)+
  theme_bw()
  # facet_wrap(~facet)
```

Poisson:

```{r}
ggplot(ggpredict(fit_p, terms = c('elevation')), aes(x, predicted)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1)+
  theme_bw()
  # facet_wrap(~facet)
```



```{r}
ITM <- read.csv('ITM.csv')
```

```{r}
all <- read.csv('all.csv')
```

```{r}
all_subset <- all[all$year_of_birth < 1930,]

all_subset$norm_yb <- scale(all_subset$year_of_birth)
all_subset$norm_el <- scale(all_subset$elevation)
all_subset$norm_itm <- scale(all_subset$number.of.lang.strat)
all_subset$norm_pop <- scale(all_subset$language.population)

all$norm_yb <- scale(all$year_of_birth)
all$norm_el <- scale(all$elevation)
all$norm_itm <- scale(all$number.of.lang.strat)
all$norm_pop <- scale(all$language.population)
all$norm_vil_pop <- scale(all$village.population)

pois_2 <- glmer(number.of.lang.strat ~ norm_el + (1|norm_yb:sex) + (1|residence:mother.tongue), data=all_subset, family='poisson')
summary(pois_2)
```

Garik M. model:

```{r}
garik_model <- glmer(number.of.lang.strat ~ sex + norm_pop + (1|norm_yb) + (1|residence), data=all, family='poisson')
summary(garik_model)
```

Normalized elevation vs. normalized village population.

```{r}
ggplot(data=all, aes(x=norm_el, y=norm_vil_pop))+
  geom_point()
```


```{r}
pois_2_pop <- glmer(number.of.lang ~ norm_el + (1|norm_yb) + (1|sex) + (1|residence) + norm_pop + norm_vil_pop, data=all, family='poisson')
summary(pois_2_pop)
```


```{r}
ITM$norm_yb <- scale(ITM$year_of_birth)
ITM$norm_el <- scale(ITM$elevation)
ITM$norm_itm <- scale(ITM$number.of.lang.strat)
```



```{r}
pois_1 <- glmer(number.of.lang.strat ~ norm_el + (1|norm_yb:sex) + (1|residence:mother.tongue), data=ITM, family='poisson')
summary(pois_1)
```

### Poisson regression vs. real data

```{r}
pois_1 <- glmer(number.of.lang.strat ~ norm_el + (1|norm_yb:sex) + (1|residence:mother.tongue), data=ITM, family='poisson')
summary(pois_1)
```

```{r}
AIC(pois_1)
```


```{r}
ITM_el <- dplyr::select(ITM, elevation, number.of.lang.strat) %>% 
  group_by(elevation) %>% 
  summarise(mean_itm = mean(number.of.lang.strat), count=n())

ITM_el
```

```{r}
pred <- ggpredict(pois_1, terms = c('norm_el'))
true_sequence <- seq(min(ITM$elevation), max(ITM$elevation), length.out=21)
```



```{r}
ggplot()+
  geom_point(data=ITM_el, aes(x=elevation, y=mean_itm, size=count), color='blue', alpha=.5)+
  geom_line(data=pred, aes(x=true_sequence, y=predicted), size=0.8)+
  geom_ribbon(aes(ymin = pred$conf.low, ymax = pred$conf.high, x=true_sequence), alpha = .15)+
  theme_bw()+
  labs(x = "Elevation", y='Average ITM', size=' Number\n of observations')

ggsave('poison_observ.jpg', width = 7, height = 4)
```

### Some geospatial ideas

#### ITM

```{r}
itm_coors <- read.csv('itm_coords_no_Hinuq.csv')
itm_coors <- itm_coors[complete.cases(itm_coors), ]
itm_coors
```

```{r}
itm_vill <- itm_coors %>% 
  group_by(index, Lat, Lon, elevation) %>% 
  summarise(mean_itm = mean(number.of.lang.strat), count=n())

itm_vill
```


```{r}
matrix <- distm(itm_vill[,c('Lon','Lat')], itm_vill[,c('Lon','Lat')], fun=distVincentyEllipsoid) / 1000
matrix[matrix < 19] <- 0
mode(matrix) <- "numeric"
rownames(matrix) <- itm_vill$index
colnames(matrix) <- itm_vill$index
net <- graph.adjacency(matrix, mode = "undirected", weighted = TRUE, diag = FALSE)

itm_vill$knn_dist <- knn(net)$knn

itm_vill$degree <- degree(net)

ggplot(itm_vill, aes(x=factor(degree), y=mean_itm, label=index))+
  theme_bw()+
  geom_boxplot()

itm_vill %>% 
  group_by(degree)  %>% 
  summarise(mean = mean(mean_itm), sd = sd(mean_itm),
            n = n())  %>% 
  mutate(se = sd / sqrt(n),
         lower.ci = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper.ci = mean + qt(1 - (0.05 / 2), n - 1) * se)  %>% 
  ggplot(aes(x=factor(degree), y=mean))+
    geom_point()+
    geom_errorbar(aes(ymin=lower.ci, ymax=upper.ci))
```


```{r}
itm_vill %>% 
  group_by(degree)  %>% 
  summarise(mean = mean(mean_itm), sd = sd(mean_itm),
            n = n())  %>% 
  mutate(se = sd / sqrt(n),
         lower.ci = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper.ci = mean + qt(1 - (0.05 / 2), n - 1) * se)  %>% 
  ggplot(aes(x=factor(degree), y=mean))+
    geom_point()+
    geom_errorbar(aes(ymin=lower.ci, ymax=upper.ci))
```


```{r}
matrix <- distm(itm_vill[,c('Lon','Lat')], itm_vill[,c('Lon','Lat')], fun=distVincentyEllipsoid) / 1000
matrix[matrix > 12] <- 0
mode(matrix) <- "numeric"
rownames(matrix) <- itm_vill$index
colnames(matrix) <- itm_vill$index
net <- graph.adjacency(matrix, mode = "undirected", weighted = TRUE, diag = FALSE)

net %>%  
  ggraph(layout = 'kk')+
  geom_edge_link(color = "orange")+
  theme_void()+
  geom_node_point(aes(size=itm_vill$mean_itm, color=itm_vill$elevation)) +  
  geom_node_text(size = 4,
                 aes(label = name), repel=TRUE)+
  labs(size='Mean ITM', color='Elevation')

ggsave('village_graph.jpg', width = 7, height = 4)
```

```{r}
matrix <- distm(itm_vill[,c('Lon','Lat')], itm_vill[,c('Lon','Lat')], fun=distVincentyEllipsoid) / 1000
matrix[matrix > 12] <- 0
mode(matrix) <- "numeric"
rownames(matrix) <- itm_vill$index
colnames(matrix) <- itm_vill$index
net <- graph.adjacency(matrix, mode = "undirected", weighted = TRUE, diag = FALSE)

ggraph(graph=net, x=itm_vill$Lon, y=itm_vill$Lat)+
  theme_void()+
  geom_edge_link(color = "orange")+
  geom_node_point(aes(color=itm_vill$mean_itm)) +  
  geom_node_text(size = 4,
                 aes(label = name), repel=TRUE)+
  labs(size='Mean ITM', color='Mean ITM')+
  annotation_map(map_data("world"), fill = NA, colour = "grey50")

ggsave('village_graph_map.jpg', width = 7, height = 4)
```



#### Families

```{r}
data_short <- data
matrix_fam <- distm(data_short[,c('Lon','Lat')], data_short[,c('Lon','Lat')], fun=distVincentyEllipsoid) / 1000
```

```{r}
matrix_fam[matrix_fam > 6] <- 0
rownames(matrix_fam) <- data_short$eng_vil_name
colnames(matrix_fam) <- data_short$eng_vil_name
net_fam <- graph.adjacency(matrix_fam, mode = "undirected", weighted = TRUE, diag = FALSE)

net_fam %>%
  # ggraph(layout = 'kk')+
  ggraph(x=data_short$Lon, y=data_short$Lat)+
  geom_edge_link(color = "orange")+
  theme_void()+
  geom_node_point(aes(color=data_short$Family), size=0.5)+
  labs(color='Family')
  # geom_node_text(aes(label = name), repel=TRUE)+
  # ggsave('village_graph_map.pdf', width = 30, height = 30)
```

```{r}
data$neighbours_6 <- degree(net_fam)
data$log_c <- log(data$census_1926)

data$elevation_round <- plyr::round_any(data$elevation, 100, f = ceiling)

data %>%
  group_by(elevation_round) %>%
  summarise(neighbours = mean(neighbours_6), n_6 = neighbours_6, size=n(), pop = census_1926) %>%
  ggplot(aes(x=elevation_round, y=neighbours))+
    geom_line()+
    geom_point(aes(y=n_6, size=pop), alpha=0.1, color='blue')+
    theme_bw()
```


```{r}
by_lang <- data  %>%  
  group_by(language, Family)  %>%  
  summarise(Lon = mean(Lon), Lat = mean(Lat), n = n())

m_l <- distm(by_lang[,c('Lon','Lat')], by_lang[,c('Lon','Lat')], fun=distVincentyEllipsoid) / 1000
m_l[m_l > 25] <- 0
rownames(m_l) <- by_lang$language
colnames(m_l) <- by_lang$language
net_l <- graph.adjacency(m_l, mode = "undirected", weighted = TRUE, diag = FALSE)

net_l %>%
  ggraph(layout = 'kk')+
  geom_edge_link(color = "orange")+
  theme_void()+
  geom_node_point(aes(color=by_lang$Family))+
  geom_node_text(size = 2,
                aes(label = name), repel=TRUE)+
  labs(color='Family')
```

How many members of different families are around each village:

```{r}
data_short <- data
matrix_fam <- distm(data_short[,c('Lon','Lat')], data_short[,c('Lon','Lat')], fun=distVincentyEllipsoid) / 1000
matrix_fam[matrix_fam > 100] <- 0
net_fam <- graph.adjacency(matrix_fam, mode = "undirected", weighted = TRUE, diag = FALSE)

V(net_fam)$label <- data$Family

SameLabel <-  function(e) {
    V(net_fam)[ends(net_fam, e)[1]]$label == V(net_fam)[ends(net_fam, e)[2]]$label }
g2 <- delete_edges(net_fam, which(sapply(E(net_fam), SameLabel)))

data_short$degree <- degree(g2)

ggplot(data=data_short, aes(x=Family, y=degree))+
  geom_boxplot()
  # geom_dotplot(aes(fill=Family, color=Family), binaxis = "y", stackdir = "center", position = "dodge", binwidth = 0.45)

# ggsave('degree_family.jpg', width = 7, height = 4)
```



```{r}
data_short  %>% 
  group_by(Family)  %>% 
  summarise(mean = mean(degree))
```

```{r}
data_itm_fam <- read.csv('itm_coords_Family.csv')
```

```{r}
itm_vill_fam <- data_itm_fam %>% 
  group_by(index, Lat, Lon, elevation, Family, mother.tongue) %>% 
  summarise(mean_itm = mean(number.of.lang.strat), count=n())

itm_vill_fam <- itm_vill_fam[!itm_vill_fam$index == 'Hinuq',]
itm_vill_fam[itm_vill_fam$mother.tongue == 'кадарский даргинский',]$Family <- 'Dargwa'
itm_vill_fam[itm_vill_fam$mother.tongue == 'сирхинский даргинский',]$Family <- 'Dargwa'
itm_vill_fam[itm_vill_fam$mother.tongue == 'цудахарский даргинский',]$Family <- 'Dargwa'
itm_vill_fam[itm_vill_fam$mother.tongue == 'гдымско-фийский лезгинский',]$Family <- 'Lezgic'
itm_vill_fam[itm_vill_fam$mother.tongue == 'цахурский',]$Family <- 'Lezgic'
itm_vill_fam[itm_vill_fam$mother.tongue == 'азербайджанский',]$Family <- 'Turkic'
itm_vill_fam[itm_vill_fam$mother.tongue == 'ахвахский',]$Family <- 'Andic'
itm_vill_fam$Lat <- itm_vill$Lat
itm_vill_fam$Lon <- itm_vill$Lon

itm_vill_fam
```


```{r}
matrix <- distm(itm_vill_fam[,c('Lon','Lat')], itm_vill_fam[,c('Lon','Lat')], fun=distVincentyEllipsoid) / 1000
matrix[matrix > 25] <- 0
mode(matrix) <- "numeric"
rownames(matrix) <- itm_vill_fam$index
colnames(matrix) <- itm_vill_fam$index
net <- graph.adjacency(matrix, mode = "undirected", weighted = TRUE, diag = FALSE)
# V(net)$label <- itm_vill_fam$Family
V(net)$label <- itm_vill_fam$mother.tongue
SameLabel <-  function(e) {
    V(net)[ends(net, e)[1]]$label == V(net)[ends(net, e)[2]]$label }
g2 <- delete_edges(net, which(sapply(E(net), SameLabel)))

itm_vill_fam$deg <- degree(g2)

# DiffLabel <-  function(e) {
#     V(net)[ends(net, e)[1]]$label != V(net)[ends(net, e)[2]]$label }
# g3 <- delete_edges(net, which(sapply(E(net), DiffLabel)))
# 
itm_vill_fam$deg <- degree(g2)
# itm_vill_fam$deg_tot <- degree(g3)
# itm_vill_fam$deg <- log(itm_vill_fam$deg_n/itm_vill_fam$deg_tot)

g2 %>%  
  ggraph(layout = 'kk')+
  geom_edge_link(color = "orange", alpha=0.5)+
  theme_void()+
  geom_node_point(aes(color=itm_vill_fam$Family))+
  scale_color_brewer(palette="Dark2")+
  labs(color='Family')

# ggsave('village_graph.jpg', width = 7, height = 4)
```

```{r}
ggplot(itm_vill_fam, aes(y=mean_itm, x=deg))+
  geom_point()
```


```{r}
ggplot(itm_vill_fam, aes(x=Family, y=deg))+
  geom_boxplot()+
  labs(x='Degree')
```




```{r}
itm_deg_m <- lmer(mean_itm ~ deg + (1|Family:mother.tongue) + count, itm_vill_fam)
summary(itm_deg_m)
```


```{r}
pred <- ggpredict(itm_deg_m, terms = c('deg'))

ggplot()+
  geom_point(data=itm_vill_fam, aes(y=mean_itm, x=deg, color=Family))+
  geom_text_repel(data=itm_vill_fam, aes(label=index, y=mean_itm, x=deg))+
  geom_line(data=pred, aes(x=pred$x, y=predicted), size=0.8)+
  geom_ribbon(aes(ymin = pred$conf.low, ymax = pred$conf.high, x=pred$x), alpha = .15)+
  scale_color_brewer(palette="Dark2")
```

Data from bigger dataset:

```{r}
overlap <- data_short[data_short$eng_vil_name %in% itm_vill_fam$index,]$eng_vil_name
ov_itm_vill_fam <- itm_vill_fam[itm_vill_fam$index %in% overlap,]
data_all <- data_short[data_short$eng_vil_name %in% itm_vill_fam$index,]
itm_vill_fam_m <- merge(x= ov_itm_vill_fam, y = data_all[, c('eng_vil_name', 'degree')], by.x='index', by.y = 'eng_vil_name', all=TRUE, copy=TRUE)

itm_deg <- lmer(mean_itm ~ degree + (1|Family:mother.tongue) + count, itm_vill_fam_m)
summary(itm_deg)

pred <- ggpredict(itm_deg, terms = c('degree'))

ggplot()+
  geom_point(data=itm_vill_fam_m, aes(y=mean_itm, x=degree, color=Family))+
  geom_text_repel(data=itm_vill_fam_m, aes(label=index, y=mean_itm, x=degree))+
  geom_line(data=pred, aes(x=pred$x, y=predicted), size=0.8)+
  geom_ribbon(aes(ymin = pred$conf.low, ymax = pred$conf.high, x=pred$x), alpha = .15)+
  scale_color_brewer(palette="Dark2")
```